SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization

### 2013-01-06 paper/paper.tex
paper: implementation and usage.

\subsection{Automated Precondition Strengthening}
The extrapolation algorithm generalizes a counterexample, but in the
``neighborhood'' of he original counterexample.  In particular, all
generalizations are from the same variant as the original counterexample.  To
help the programmer in the generalization step, we would also like a way to test
the property again, ensuring we get counterexamples (if they exist) outside of
the neighborhood of the original one.

@@

\section{Implementation}\label{sec:implementation}

\todo{generics}

\todo{GADTs? don't work}
\todo{add lazy testing---proof value is unused}

@@

%-------------------------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}

\todo{Use Bayes' rule to do extrapolation?}

@@

>>>>

\subsection{Automated Precondition Strengthening}\label{sec:precondition}
The extrapolation algorithm generalizes a counterexample, but in the
``neighborhood'' of he original counterexample.  In particular, all
generalizations are from the same variant as the original counterexample.  To
help the programmer in the generalization step, we would also like a way to test
the property again, ensuring we get counterexamples (if they exist) outside of
the neighborhood of the original one.

@@

\section{Implementation and Usage}\label{sec:implementation}

The implementation of SmartCheck is written in Haskell and is designed to test
Haskell programs.  The source code is licensed BSD3 and is freely
available.\footnote{\url{https://github.com/leepike/SmartCheck.git}}

SmartCheck generically operates over arbitrary algebraic data and so uses a
generics library to encode generic traversals.  Specifically, SmartCheck uses
``GHC Generics'', a recently-developed generics library for
Haskell~\cite{generics}.  The generics library allows algebraic data types to
automatically derive the type class \ttp{SubTypes} presented in
Section~\ref{sec:preliminaries}.  One limitation with GHC Generics is that it
does not (currently) support generalized algebraic data types~\cite{gadts}.

The data type to be tested must derive the \ttp{Read}, \ttp{Show},
\ttp{Typeable}, and \ttp{Generic} type classes.  Deriving \ttp{Typeable} and
\ttp{Generic} require using the GHC language extensions \ttp{DeriveDataTypeable}
and \ttp{DeriveGeneric}, respectively.  \ttp{Typeable} is used for dynamic
typing in defining the \ttp{replace} method of \ttp{SubTypes} since it is
unknown at compile-time whether the value to be replaced and replacing value
have the same types.

Then, the user simply declares, for a data type \ttp{D},
%
\begin{code}
instance Subtypes D
\end{code}
%

SmartCheck does not implement a counterexample discovery algorithm.  While any
method can be used to discover an initial counterexample, SmartCheck by default
uses QuickCheck to discover an initial counterexample.  If QuickCheck is used,
instances for \ttp{Arbitrary} must be provided.

A property may be a function taking any number of arguments, but SmartCheck only
operates on a single input during testing, which is assumed to be the first
argument to the function.  Our experience is that often a program operates over
one large structure together with smaller inputs.  While multiple inputs could
be analyzed simultaneously by SmartCheck, it is not clear how to clearly
represent multiple large counterexamples to the user---the implementation
currently prints to standard out.  (In future work, a web-based interface could help.)

A read-eval-print loop is presented to the user, allowing her to iterate shrink
and generalize counterexamples, and then generate new counterexamples after
strengthening the property's precondition as described in
Figure~\ref{sec:precondition}.

SmartCheck is executed using
%
\begin{code}
> smartCheck args prop
\end{code}
%
\noindent
where \ttp{args} are the arguments passed in, and \ttp{prop} is the property
being tested.  The arguments is a record type that includes the values described
above to control SmartCheck's behavior.

@@

\todo{stuff from SC repo}

%-------------------------------------------------------------------------------
\section{Conclusions and Future Work}\label{sec:conclusions}

\todo{how to display counterexamples}
\todo{Use Bayes' rule to do extrapolation?}

@@

\todo{add lazy testing---proof value is unused}
